---
title: "2. Generating FVS Inputs from Treemap"
author: "Carmen Tubbesing"
date: "`r Sys.Date()`"
format: 
  html:
    embed-resources: TRUE
    toc: TRUE
---

```{r, include = F}
source("project_vars.R", local = TRUE)
```

# This run is for project `r project_ID`, AKA `r proj_name`

# Files needed:
-  Blank FVS Init excel workbook
-  Outputs from Step 1, including list of TreeMap Stand ID's

```{r, include = F}
# load libraries
library(terra) # for handling spatial data
library(dplyr) # for data manipulation
library(readxl) # for reading excel workbooks
library(sf) # for shapefiles
require(writexl) # for saving excel files
```

# Setup

### Load areas files from Script 1

These files were created by the script `1-Intersect_TreeMap_treatments.qmd`

# Individual treatment areas files 
```{r}
# list the excel files in temp_dir that end with "_areas"
area_files <- list.files(temp_dir, pattern = paste0("^", project_ID, ".*_areas\\.xlsx$"), full.names = TRUE)
if(length(area_files)>0){
  print("input files with areas of each StandID are:")
  print(tools::file_path_sans_ext(basename(area_files)))
} else{
  print("***ALERT*** Missing input files with areas of each StandID ")
}
```

# Individual impact area files 
```{r}
# list the excel files in inputs_dir that end with "_areas_TM"
IA_files <- list.files(temp_dir, pattern = paste0("^IA_areas_", project_ID, "_.*"), full.names = TRUE)
if(length(IA_files)>0){
  print("Impact Area input files with areas of each StandID are:")
  print(tools::file_path_sans_ext(basename(IA_files)))
} else{
  print("***ALERT*** Missing Impact Area input files with areas of each StandID ")
}
```


## Load TreeMap tree tables
```{r}
trees16file <- file.path(treemap_dir, "TreeMap2016/TreeMap2016_tree_table.csv")
trees20file <- file.path(treemap_dir, "TreeMap2016/TreeMap2016_tree_table.csv")
trees22file <- file.path(treemap_dir, "TreeMap2022/TreeMap2022_CONUS_Tree_Table.csv")
```

Only load the TM vintage you need because it is slow to load
```{r}
if(tm == 2016){
  trees16 <- read.csv(trees16file, colClasses = c("character", "character"))
  tm_trees <- trees16 %>% 
    rename(TM_ID = tm_id)
} else if(tm == 2020){
  trees20 <- read.csv(file = trees20file, colClasses = c("character", "character"))
  tm_trees <- trees20 %>% 
    rename(TM_ID = tm_id)
} else if(tm == 2022){
  trees22 <-  read.csv(file = trees22file, colClasses = c("character", "character"))
  tm_trees <- trees22
} else{
  print("ERROR: Not a valid TreeMap year")
}
```


## Prep tm_trees  
```{r}
tm_trees <- tm_trees %>% 
  mutate(TPA_UNADJ = as.numeric(TPA_UNADJ)) %>%  
  mutate(DIA = as.numeric(DIA))
```

## Remove trees without DIA and Ht values
```{r}
tm_trees <- tm_trees %>% 
  filter(!is.na(DIA) & !is.na(HT))
```

## Remove records where TPA_UNADJ is NA
```{r}
tm_trees <- tm_trees %>% 
  filter(!is.na(TPA_UNADJ))
```

## Read in blank FVS Init .xlsx sheets as blank stand init, plot init and tree init dataframes
```{r}
blank_stand <- read_excel(blank_db, sheet = "FVS_StandInit")
blank_tree <- read_excel(blank_db, sheet = "FVS_TreeInit")
```

# Function to make the new stand_init table
```{r}
# tm <- area_files[1] # for debugging
tm_stand_init <- function(area_file, blank_stand_init, proj_name, tm) {
  #' @param tm filepath for the list of treemap IDs generated in step 1, with FVS variant column
  #' @param blank_stand_init the blank stand init
  #' @param proj_name the project name
  #' @param tm the vintage of the TreeMap data you are using
  #' @return A new , filled stand init table for the AOI
  df <- read_excel(area_file) %>% 
    mutate(TM_ID = as.character(TM_ID))
  num_polygons <- length(df) #get the number of unique treemapper IDs
  variant_list <- list(unique(df$FVSVariant))
  stand_init <- blank_stand_init %>% 
    mutate(STAND_ID = as.character(STAND_ID))
  stand_init <- add_row(stand_init, STAND_ID = df$TM_ID)
  stand_init$VARIANT <- df$FVSVariant #fill variant column
  stand_init$INV_YEAR <- tm #fill inventory year column
  stand_init$GROUPS <- "All_Stands" #fill groups column
  stand_init$PROJECT_NAME <- proj_name #fill project name column
  stand_init$INV_PLOT_SIZE <- 1 #fill inventory plot size column
  stand_init$BRK_DBH <- 999 #fill brk_dbh column
  stand_init$NUM_PLOTS <- 1 #fill num_plots column
  return(stand_init)
}
```

# Function to make new tree init table

## List the columns that were generated by the LEMMA workflow
```{r}
tree_cols <- c("Stand_CN", "Stand_ID","Plot_ID","Tree_ID", "Tree_Count"  ,"History", "Species",    "DBH","DG","Ht", "HTG",    "HtTopK"  ,"CrRatio",    "Damage1"  ,    "Severity1"   , "Damage2"  ,    "Severity2"  ,  "Damage3"  ,   "Severity3" ,   "TreeValue"  ,  "Prescription", "Age"   )
```

## Tree init function
```{r}
tm_tree_init <- function(area_file, blank_tree_init, tm_tree_list) {
    df <- read_excel(area_file) %>% 
      mutate(TM_ID = as.character(TM_ID))
    df_trees_full <- merge(df, tm_tree_list, by = "TM_ID", all.x = TRUE) 
    
    if ("CN" %in% names(df_trees_full)) {
      # Filter records where 'CN' is not NA
      df_trees_full <- df_trees_full[!is.na(df_trees_full$CN), ] %>% 
      rename(Stand_CN = CN)
    } else if ("PLT_CN" %in% names(df_trees_full)) {
      df_trees_full <- df_trees_full %>% 
        rename(Stand_CN = PLT_CN)
    }
    
    df_trees <- df_trees_full %>% 
      mutate(Stand_ID = TM_ID) %>% 
      select(Stand_CN, Stand_ID, TPA_UNADJ, STATUSCD, SPCD, DIA, HT, CR) 
    
    tree_list <- df_trees %>% 
      mutate(Tree_ID = c(1:nrow(df_trees))) %>% 
      mutate(Plot_ID = 1) %>% 
      mutate(History = case_when(
        STATUSCD == "2" ~ "8",
        TRUE ~STATUSCD  # Keep existing HISTORY values if STATUSCD is not 2
      )) %>% 
      rename(Species = SPCD, 
             DBH = DIA,
             Ht = HT,
             Tree_Count = TPA_UNADJ, 
             CrRatio = CR) 
    
    missing_cols <- setdiff(tree_cols, colnames(tree_list))

    for (col in missing_cols) {
      tree_list[[col]] <- NA
    }

    # Reorder columns so they match TreeMap input format
    tree_list <- tree_list %>%
      select(all_of(tree_cols))
    
    return(tree_list)
}
```

# Build FVS inputs for treatment areas
```{r}
for (i in 1:length(area_files)){
  print(i)
  stand_table <- tm_stand_init(area_file = area_files[i], 
                blank_stand_init = blank_stand, 
                proj_name = proj_name, 
                tm = tm) 
  tree_table <- tm_tree_init(area_file = area_files[i],
             blank_tree_init = blank_tree,
             tm_tree_list = tm_trees)
  basename <- tools::file_path_sans_ext(basename(area_files[i]))
  basename <- gsub("_areas", "", basename)
  file <- paste0(fvs_input_dir, "/FVS_input_", basename, ".xlsx")
  write_xlsx(list(FVS_StandInit=stand_table, FVS_TreeInit=tree_table), file)
  print(paste("Created FVS inputs saved as", file))
}
```

# Build FVS inputs for impact areas
```{r}
if(length(IA_files)>0){
  for (i in 1:length(IA_files)){
    print(i)
    stand_table <- tm_stand_init(area_file = IA_files[i], 
                  blank_stand_init = blank_stand, 
                  proj_name = proj_name, 
                  tm = tm) 
    tree_table <- tm_tree_init(area_file = IA_files[i],
               blank_tree_init = blank_tree,
               tm_tree_list = tm_trees)
    basename <- tools::file_path_sans_ext(basename(IA_files[i]))
    basename <- gsub("areas_", "", basename)
    file <- paste0(fvs_input_dir, "/FVS_input_", basename, ".xlsx")
    write_xlsx(list(FVS_StandInit=stand_table, FVS_TreeInit=tree_table), file)
    print(paste("Created FVS inputs saved as", file))
  }
}
```

