---
title: "3. Calculate carbon results from FVS output - Fuels Reduction"
author: "Carmen Tubbesing"
date: "`r Sys.Date()`"
format: 
  html:
    embed-resources: TRUE
    toc: TRUE
---

```{r, include = F}
source("project_vars.R", local = TRUE)
```

```{r, include = F}
source("treatment_vars.R", local = TRUE)
```

```{r, include = F}
require(dplyr)
require(readxl)
require(ggplot2)
require(flextable)
```

## Project = `r proj_name`
## Project ID = `r project_ID`
## TCN = `r TCN`
## Type = `r type`

## Load the FVS output files
```{r}
if(type == "TA"){
  fvs_output_file <- paste0(project_ID, "/", project_ID, "_", TCN, "_treatment", "_carbon")
  fvs_output_file <- paste0(fvs_output_dir, fvs_output_file, ".xlsx")
  fvs_output_file
}
```

```{r}
if(type == "IA"){
  fvs_output_file <- paste0(project_ID, "/", project_ID, "_", TCN, "_IA", "_TRIAADS")
  fvs_output_file <- paste0(fvs_output_dir, fvs_output_file, ".xlsx")
  fvs_output_file
}
```

```{r}
df <- read_excel(fvs_output_file)
```

# Plot carbon over time for all stands, all runs 

This is similar to what FVS shows if you choose "wait for run." Code written by Jason Singh with help from copilot.
```{r}
df <- df %>%
  mutate(Total_Carbon = Aboveground_Total_Live + Belowground_Live)
# Plot total carbon
ggplot(df, aes(x = Year, y = Total_Carbon, color = as.factor(StandID), group = StandID)) +
  geom_line() +
  facet_wrap(~ MgmtID, ncol = 2) +
  labs(x = "Year",
       y = "Total Carbon",
       color = "StandID") +
  scale_color_manual(values = rainbow(length(unique(df$StandID)))) +
  theme_light() +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90, hjust = 1),
        strip.text = element_text(color = "#000000"),
        axis.text = element_text(color = "#000000"))

# Save output file
ggsave("Total_C_Graph.png", width = 6, height = 6, units = "in")
```

## Load the data tables with areas
```{r}
if(type == "TA"){
  areas_file <- paste(project_ID, TCN, "areas", sep = "_")
  areas_file <- paste0(temp_dir, "/", areas_file, ".xlsx")
  areas <- read_excel(areas_file)
}
```

```{r}
if(type == "IA"){
  areas_file <- paste("IA", "areas", project_ID, TCN, sep = "_")
  areas_file <- paste0(temp_dir, "/", areas_file, ".xlsx")
  areas <- read_excel(areas_file)
}
```

# Delete non-forested stands & mutate stand ID to match FVS output format
```{r}
areas <- areas %>% 
  filter(TM_ID > 0) %>% 
  mutate(Stand_ID = as.character(TM_ID))
```


# Plot average aboveground live values
```{r}
df_means <- df %>% 
  group_by(Year, MgmtID) %>% 
  summarize(Aboveground_Total_Live = mean(Aboveground_Total_Live)) 
df_means <- df_means %>% 
  ungroup()
```



```{r}
ggplot(df_means) +
  geom_line(aes(x = Year, y = Aboveground_Total_Live, col = MgmtID, group = MgmtID)) +
  labs(
    y = "Live Aboveground Carbon (C)") +
  theme_minimal()+
  ggtitle("AVERAGE AGL carbon across stands (not weighted by area of each stand)")
```

# Plot by treatment and stand, carbon over time
```{r}
if("BSWF" %in% df$MgmtID){
  ggplot(df %>% filter(MgmtID == "BSWF"))+
    geom_line(aes(x = Year, y = Aboveground_Total_Live, col = StandID, group = StandID))+
    labs(
      y = "Live Aboveground Carbon (C)") +
    
    ggtitle("BSWF")
    
  ggplot(df %>% filter(MgmtID == "TRWF"))+
    geom_line(aes(x = Year, y = Aboveground_Total_Live, col = StandID, group = StandID))+
    labs(
      y = "Live Aboveground Carbon (C)") +
    
    ggtitle("TRWF")
    
  
  ggplot(df %>% filter(MgmtID == "TRNF"))+
    geom_line(aes(x = Year, y = Aboveground_Total_Live, col = StandID, group = StandID))+
    labs(
      y = "Live Aboveground Carbon (C)") +
    
    ggtitle("TRNF")

}
```

```{r}
if("BSPM" %in% df$MgmtID){
    ggplot(df %>% filter(MgmtID == "BSPM"))+
      geom_line(aes(x = Year, y = Aboveground_Total_Live, col = StandID, group = StandID))+
      labs(
        y = "Live Aboveground Carbon (C)") +
      
      ggtitle("BSPM")
  
    ggplot(df %>% filter(MgmtID == "TRPM"))+
      geom_line(aes(x = Year, y = Aboveground_Total_Live, col = StandID, group = StandID))+
      labs(
        y = "Live Aboveground Carbon (C)") +
      
      ggtitle("TRPM")
}
```

```{r}
if("IAPM" %in% df$MgmtID){
    ggplot(df)+
      geom_line(aes(x = Year, y = Aboveground_Total_Live, col = StandID, group = StandID))+
      labs(
        y = "Live Aboveground Carbon (C)") +
      
      ggtitle("IAPM")
}
```

# Plot the distribution of ending carbon in each scenario
```{r}
ggplot(df %>%  filter(Year == end_year), aes(x = factor(MgmtID), y = Aboveground_Total_Live)) +
  geom_violin(fill = "lightblue", color = "black", trim = FALSE) +
  geom_boxplot(width = 0.1, outlier.shape = NA, alpha = 0.5) +  # optional boxplot overlay
  labs(
    x = "Management ID",
    y = "Live Aboveground Carbon (C) at project end",
    title = "Distribution of Total_C by MgmtID"
  ) +
  theme_minimal()
```


# Merge with areas from pre-FVS files and calculate results
-  Calculate total ending year carbon for each scenario
-  Print results in the order they appear in `FRM_Calculator_Tool`


```{r}
df_list <-  unique(df$MgmtID)
```

```{r}
for (name in df_list) {
  df_i <- df %>% 
    filter(MgmtID == name)
   
  df_i <- left_join(df_i, areas, join_by(StandID==Stand_ID))
  
  if (any(is.na(df_i$TM_ID))) {
    message("ERROR: The TM_ID column in the FVS output file and the areas file don't match.")
    message("Are you using the right areas file?")
    next()
  }

  # Create a new column, e.g. "Biomass_Area"
  df_carbon <- df_i%>% 
    filter(Year == end_year) %>% 
    mutate(Total_C = (Aboveground_Total_Live + Belowground_Live) * area_ac)
  
  # Save it back into the list
  assign(name, df_carbon)
  
  # Sum carbon and label it
  sum <- sum(df_carbon$Total_C)
 
  print(paste("total C in", name, "=", round(sum, 0), "MTC"))
  
  assign(paste(name, "sum_C", sep = "_"), sum)
  
  if(("TRNF"==name| "TRPM" ==name) & "LIVECRM" %in% names(df_i)){
    df_i <- df_i %>% 
      mutate(livecrem_MTC = LIVECREM*area_ac)    
    livecrem_MTC <- sum(na.omit(df_i$livecrem_MTC))
    print(paste("Carbon removed (MTC) =", round(livecrem_MTC,2)))
    removed_BDT <- livecrem_MTC/0.50/0.907185
    print(paste("Biomass removed via mechanical treatments (BDT) =", round(removed_BDT,2)))
  } else if(("TRNF"==name| "TRPM" ==name) & !("LIVECRM" %in% names(df_i))){
  print("no biomass removed via mechanical treatments ('BR')")
}
  
}
```

# Print tables for treatment area carbon by stand ID
```{r}
if(type=="TA"){
   TRNF <- TRNF %>% 
    mutate(Carbon_per_ac_TRNF = round(Total_Carbon,2), area_ac = round(area_ac,2), Total_C_TRNF = round(Total_C,2)) %>% 
    select(StandID, Carbon_per_ac_TRNF, area_ac, Total_C_TRNF)
  
   TRWF <- TRWF %>% 
    mutate(Carbon_per_ac_TRWF = round(Total_Carbon,2), area_ac = round(area_ac,2), Total_C_TRWF = round(Total_C,2)) %>% 
    select(StandID, Carbon_per_ac_TRWF, area_ac, Total_C_TRWF)

   TRNF_TRWF <- full_join(TRNF, TRWF) %>% 
     select(StandID, area_ac, Carbon_per_ac_TRNF, Carbon_per_ac_TRWF, Total_C_TRNF, Total_C_TRWF)
  flextable(TRNF_TRWF) 
}
```

