---
title: "3. Calculate carbon results from FVS output - Fuels Reduction"
author: "Carmen Tubbesing"
date: "`r Sys.Date()`"
format: 
  html:
    embed-resources: TRUE
    toc: TRUE
---

# Setup
```{r, include = F}
require(dplyr)
require(readxl)
require(ggplot2)
```

# This run is for project `Hughes CFIP`
## Treatment = `8GG25301`
## Type = `IA`


# Define folders and variables 
*Note:* Replace the values in this section for the file folder on your computer

```{r}
end_year <- 2077
```


```{r}
fvs_output_dir = "~/../Data_heavy_projects/Forest_QM_automation/FVS_Output/"
areas_dir <- "~/../Data_heavy_projects/Forest_QM_automation/temp"
```

## Load the FVS output files
```{r}
df <- read_excel(paste0(fvs_output_dir, "/8GG25301_IA_carbon_TRIAADS.xlsx"))
```

## Load the data tables with areas
```{r}
areas <- read_excel(paste0(areas_dir, "/IA_areas_8GG25301_1new.xlsx"))
```


# Delete non-forested stands & mutate stand ID to match FVS output format
```{r}
areas <- areas %>% 
  filter(TM_ID > 0) %>% 
  mutate(Stand_ID = as.character(TM_ID))
```

```{r}
df_means <- df %>% 
  group_by(Year, MgmtID) %>% 
  summarize(Aboveground_Total_Live = mean(Aboveground_Total_Live)) 
df_means <- df_means %>% 
  ungroup()
```


```{r}
ggplot(df_means) +
  geom_line(aes(x = Year, y = Aboveground_Total_Live, col = MgmtID, group = MgmtID)) +
  labs(
    y = "Live Aboveground Carbon (C)") +
  theme_minimal()
```

# Plot by treatment and stand, carbon over time
```{r}
if("BSWF" %in% df$MgmtID){
  ggplot(df %>% filter(MgmtID == "BSWF"))+
    geom_line(aes(x = Year, y = Aboveground_Total_Live, col = StandID, group = StandID))+
    labs(
      y = "Live Aboveground Carbon (C)") +
    
    ggtitle("BSWF")
    
  ggplot(df %>% filter(MgmtID == "TRWF"))+
    geom_line(aes(x = Year, y = Aboveground_Total_Live, col = StandID, group = StandID))+
    labs(
      y = "Live Aboveground Carbon (C)") +
    
    ggtitle("TRWF")
    
  
  ggplot(df %>% filter(MgmtID == "TRNF"))+
    geom_line(aes(x = Year, y = Aboveground_Total_Live, col = StandID, group = StandID))+
    labs(
      y = "Live Aboveground Carbon (C)") +
    
    ggtitle("TRNF")

}
```

```{r}
if("BSPM" %in% df$MgmtID){
    ggplot(df %>% filter(MgmtID == "BSPM"))+
      geom_line(aes(x = Year, y = Aboveground_Total_Live, col = StandID, group = StandID))+
      labs(
        y = "Live Aboveground Carbon (C)") +
      
      ggtitle("BSPM")
  
    ggplot(df %>% filter(MgmtID == "TRPM"))+
      geom_line(aes(x = Year, y = Aboveground_Total_Live, col = StandID, group = StandID))+
      labs(
        y = "Live Aboveground Carbon (C)") +
      
      ggtitle("TRPM")
}
```

```{r}
if("IAPM" %in% df$MgmtID){
    ggplot(df)+
      geom_line(aes(x = Year, y = Aboveground_Total_Live, col = StandID, group = StandID))+
      labs(
        y = "Live Aboveground Carbon (C)") +
      
      ggtitle("IAPM")
}
```

# Plot the distribution of ending carbon in each scenario
```{r}
ggplot(df %>%  filter(Year == end_year), aes(x = factor(MgmtID), y = Aboveground_Total_Live)) +
  geom_violin(fill = "lightblue", color = "black", trim = FALSE) +
  geom_boxplot(width = 0.1, outlier.shape = NA, alpha = 0.5) +  # optional boxplot overlay
  labs(
    x = "Management ID",
    y = "Live Aboveground Carbon (C) at project end",
    title = "Distribution of Total_C by MgmtID"
  ) +
  theme_minimal()
```


# Merge with areas from pre-FVS files and calculate results
-  Calculate total ending year carbon for each scenario
-  Print results in the order they appear in `FRM_Calculator_Tool`


```{r}
df_list <-  unique(df$MgmtID)
```

```{r}

# Loop through each data frame
for (name in df_list) {
  df_i <- df %>% 
    filter(MgmtID == name)
   
  df_i <- left_join(df_i, areas, join_by(StandID==Stand_ID))
  
  # JUST FOR MARSHALL AND SEARS
  df_i <- df_i %>% 
    filter(!is.na(TM_ID))
  
  if (any(is.na(df_i$TM_ID))) {
    message("ERROR: The TM_ID column in the FVS output file and the areas file don't match.")
    message("Are you using the right areas file?")
    next()
  }

  # Create a new column, e.g. "Biomass_Area"
  df_carbon <- df_i%>% 
    filter(Year == end_year) %>% 
    mutate(Total_C = (Aboveground_Total_Live + Belowground_Live) * area_ac)
  
  # Save it back into the list
  assign(name, df_carbon)
  
  # Sum carbon and label it
  sum <- sum(df_carbon$Total_C)
 
  print(paste("total C in", name, "=", round(sum, 0), "MTC"))
  assign(paste(name, "sum_C", sep = "_"), sum)
  
  if(("TRNF"==name| "TRPM" ==name) & "LIVECRM" %in% names(df_i)){
    df_i <- df_i %>% 
      mutate(livecrem_MTC = LIVECREM*area_ac)    
    livecrem_MTC <- sum(na.omit(df_i$livecrem_MTC))
    print(paste("Carbon removed (MTC) =", round(livecrem_MTC,2)))
    removed_BDT <- livecrem_MTC/0.50/0.907185
    print(paste("Biomass removed via mechanical treatments (BDT) =", round(removed_BDT,2)))
}
  
}
```


