---
title: "Reformat form responses"
author: "Carmen Tubbesing"
date: "`r Sys.Date()`"
format: 
  html:
    embed-resources: TRUE
    toc: TRUE
---

```{r, include = F}
require(tidyverse)
require(readxl)
require(writexl)
```

# Load form responses

```{r}
df <- read_excel("C:/Users/ctubbesi/OneDrive - California Air Resources Board/Documents/CCI/QMs/Forest Health QM/online form/Forest Restoration & Management QM 2.xlsx")
```

# Define what row you want to start on
```{r}
start_id <- 165
start_id_supp <- 4
```

# Replace non-breaking spaces with normal spaces
```{r}
names(df) <- gsub("\u00A0", " ", names(df))
```

# Define questions that are in the wrong order
```{r}
fuels_acreage_1_q <-  "How many acres will be treated in this Fuels Reduction Activity 1?" 
fuels_acreage_2_q <- "How many acres will be treated in Fuels Reduction Activity 2?"  
fuels_acreage_3_q <- "How many acres will be treated in Fuels Reduction Activity 3?" 
fuels_acreage_4_q <- "How many acres will be treated in Fuels Reduction Activity 4?" 
fuels_site_class_1_q <- "What is the Forest Productivity Site Class in the area where Fuels Reduction Activity 1 will occur?" 
fuels_site_class_2_q <- "What is the Forest Productivity Site Class in the area where Fuels Reduction Activity 2 will occur?"
fuels_site_class_3_q <- "What is the Forest Productivity Site Class in the area where this treatment will occur?1" 
fuels_site_class_4_q <-  "What is the Forest Productivity Site Class in the area where this treatment will occur?"
```

# Find the column numbers of those columns

```{r}
fuels_acreage_1_id <- match(fuels_acreage_1_q, names(df))
fuels_acreage_1_id
fuels_acreage_2_id <- match(fuels_acreage_2_q, names(df))
fuels_acreage_2_id
fuels_acreage_3_id <- match(fuels_acreage_3_q, names(df))
fuels_acreage_3_id
fuels_acreage_4_id <- match(fuels_acreage_4_q, names(df))
fuels_acreage_4_id


site_class_1_id <- match(fuels_site_class_1_q, names(df))
site_class_1_id
site_class_2_id <- match(fuels_site_class_2_q, names(df))
site_class_2_id
site_class_3_id <- match(fuels_site_class_3_q, names(df))
site_class_3_id
site_class_4_id <- match(fuels_site_class_4_q, names(df))
site_class_4_id
```

The column number for the first fuels reduction acreage question is `r fuels_acreage_1_id`. \
The column number for the second fuels reduction acreage question is `r fuels_acreage_2_id`. \
The column number for the first site class question is `r site_class_1_id`. \
The column number for the second site class question is `r site_class_2_id`

# Identify Project ID column

```{r}
id_q<- match("Please enter the Applicant ID or Grant ID for your CCI project. ", names(df))
names(df)[id_q] <- "Project_ID_CCI"
id_q_prop4 <- match("Please enter the Applicant ID for your Prop 4 application.", names(df))
names(df)[id_q_prop4] <- "Project_ID_Prop4"
```

## Create a project ID column that's the CCI ID, unless there is no CCI ID, in which case it's the Prop 4 ID

```{r}
df <- df %>% 
  mutate(Project_ID = ifelse(Project_ID_CCI %in% c("NA", "N/A", "n/a"), Project_ID_Prop4, Project_ID_CCI))
```

```{r}
siteprep_yn <-  "Does your project include site preparation?" 
siteprep_2  <- "Does your project include additional site prep activities?" 
```

```{r}
siteprep_yn <- match(siteprep_yn, names(df))
siteprep_yn
siteprep_2 <- match(siteprep_2, names(df))
siteprep_2

```

```{r}
reforestation_yn <-  "Does your project include planting for reforestation?" 
 reforestation_2  <- "Does your project contain additional reforestation activity(ies)?1" 
reforestation_3 <- "Does your project contain additional reforestation activity(ies)?" 
```

```{r}
reforestation_yn <- match(reforestation_yn, names(df))
reforestation_yn
reforestation_2  <- match(reforestation_2 , names(df))
reforestation_2 
reforestation_3 <- match(reforestation_3, names(df))
reforestation_3
```

```{r}
pest_yn_q <- "Does your project include pest management?"
biomass_yn_q <- "Does your project include biomass utilization?"
fuels_yn_q <- "Does your project include fuels reduction treatments or broadcast burn(s)?"
fuels_2_q <- "Does your project contain additional fuel reduction treatments?"
fuels_3_q <- "Does your project contain additional fuel reduction activities?"
fuels_4_q <-"Does your project contain additional fuel reduction activities?1"
```

```{r}
pest_yn <- match(pest_yn_q, names(df))
pest_yn
biomass_yn <- match(biomass_yn_q, names(df))
biomass_yn
conservation_yn <- grep("^Does your project result in a conservation easement", names(df))
conservation_yn
fuels_yn <- match(fuels_yn_q, names(df))
fuels_yn
fuels_2 <- match(fuels_2_q, names(df))
fuels_2
fuels_3 <- match(fuels_3_q, names(df))
fuels_3
fuels_4 <- match(fuels_4_q, names(df))
fuels_4
```

# Define function to create each of the six sub-data-frames

```{r}
convert_df <- function(df, row, col_min, col_max){
    df_new <- df[row, c(col_min:col_max)] %>% t() %>% as_tibble()
  colnames(df_new) <- "Answer"
  df_new <- df_new %>% 
    mutate(Question = colnames(df[c(col_min:col_max)])) %>% 
    select(Question, Answer)
  return(df_new)
}
```

# Transpose, clean, and write an excel file for each project

```{r}
df_full <- df
df <- df %>% filter(Id>=start_id)
for(i in 1:nrow(df)){
  id <- df[i, "Project_ID"] %>% unlist()
  name <- df[i, "Project Name"] %>% unlist()
  if(is.na(id)){
    id <- name
  }
  if(sum(na.omit(df_full$Project_ID) == id) >= 2){
    print("ALERT! This project ID has already been submitted. Please delete duplicate response and start over.")
  }
  if(sum(df_full$`Project Name` == name) >= 2 ){
    print("ALERT! This project name has already been submitted. Please delete duplicate response and start over.")
  }
  print(unlist(id))
  df1 <- convert_df(df=df, row=i, col_min=1, col_max = siteprep_yn-1)
  df2 <- convert_df(df=df, row=i, siteprep_yn, pest_yn-1)
  df3 <- convert_df(df=df, row=i, pest_yn, biomass_yn-1) %>% 
    filter(Question != fuels_acreage_1_q)
  df4 <- convert_df(df=df, row=i, biomass_yn, conservation_yn-1) %>% 
    filter(Question != fuels_acreage_2_q)
  #df5 <- convert_df(df=df, row=i, conservation_yn, fuels_yn-1) %>% 
   # filter(Question != fuels_site_class_1_q)
  df6 <- convert_df(df=df, row=i, col_min=fuels_yn, col_max = fuels_2-1)
  df7 <- convert_df(df=df, row=i, fuels_2, fuels_3-1)
  df8 <- convert_df(df=df, row=i, fuels_3, fuels_4-1)
  df9 <- convert_df(df=df, row=i, fuels_4, ncol(df)-1)
  
  df6 <- bind_rows(df6, df7, df8, df9)
  
  # Remove questions where the answer is NA for all 6 dfs
  dfs <- c("df1", "df2", "df3", "df4",  "df6")
  for(j in 1:length(dfs)){
      dfj <- get(dfs[j])
      dfj <- dfj[!is.na(dfj$Answer), ]
      assign(dfs[j], dfj)
  }
  
  unnecessary_qs <- c("ID", "Start time", "Email", "Name", "Last modified time")
  df1 <- df1 %>% 
    filter(!Question %in% unnecessary_qs)
  df6_remove <- which(grepl("Does your project contain additional fuel reduction treatments?", df6$Question))
  df6 <- df6[-df6_remove,]
    df6_remove <- which(grepl("Does your project include fuels reduction treatments?", df6$Question))
  df6 <- df6[-df6_remove,]

  
  # Write to excel
  file_name <- paste0("../../form_responses/raw/", id, ".xlsx")
  sheets <- list("Project"=df1, 
                 "Reforestation" =df2, 
                 "Pest_management"=df3, 
                 "Biomass"=df4, 
                 "Fuels_Reduction"=df6)
  write_xlsx(sheets, file_name)
}
print("finished reshaping responses from main form")
```

# Repeat for supplementary data

## Load
```{r}
dfs <- read_excel("C:/Users/ctubbesi/OneDrive - California Air Resources Board/Documents/CCI/QMs/Forest Health QM/online form/Intake Form Supplementary.xlsx")
```

## Rename ID column

```{r}
head(names(dfs), n=15)
id_q<- grep("Prop 4 Forest Health", names(dfs)) 
id_q
names(dfs)[id_q] <- "Project_ID_Prop4"
```

```{r}
id_q<- grep("CCI Forest Health project",names(dfs)) 
id_q
names(dfs)[id_q] <- "Project_ID_CCI"
```

##Replace non-breaking spaces with normal spaces
```{r}
names(dfs) <- gsub("\u00A0", " ", names(dfs))
```

```{r}
dfs <- dfs %>% filter(Id>=start_id_supp)
dfs <- dfs %>% 
  mutate(Project_ID = ifelse(Project_ID_CCI %in% c("NA", "N/A", "n/a"), Project_ID_Prop4, Project_ID_CCI))
for(i in 1:nrow(dfs)){
  num <- dfs[i, "Id"] %>% unlist()
  id <- dfs[i, "Project_ID"] %>% unlist()
  name <- dfs[i, "Project Name"] %>% unlist()
  if(is.na(id)){
    id <- name
  }
  print(unlist(id))
  dfi <- dfs[i,] %>% t() %>% as_tibble()
  colnames(dfi) <- "Answer"
  
  # Remove unnecessary columns like start time and end time
  unnecessary_qs <- c("Id", "Completion time", "Start time", "Email", "Name", "Last modified time")
  dfi <- dfi %>% 
    mutate(Question = colnames(dfs)) %>% 
    filter(!Question %in% unnecessary_qs) %>% 
    select(Question, Answer)
  
  dfi_remove <- which(grepl("Does your project contain additional fuel reduction treatments?", dfi$Question))
  dfi <- dfi[-dfi_remove,]
  
  # Remove empty cells
  dfi <- dfi[!is.na(dfi$Answer), ]
  
  # Write to excel
  file_name <- paste0("../../form_responses/raw/supplementary/", id, "_", num, "_supp.xlsx")
  write_xlsx(dfi, file_name)
}
print("Finished reshaping Supplementary responses")
```
